---
title: "HW2"
format: html
editor: visual
---

# Homework 2

## Task 1 - Vectors

### Create Vectors

```{r}
bpBefore = c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)
bpAfter = c( 114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)
```

### Assign Names to Each Vector

```{r}
names = paste("Subject", 1:20, sep = "_")

names(bpBefore) = names
names(bpAfter) = names

bpAfter[10]
```

### Calculate Change in Blood Pressure for Each Patient

```{r}
bpDelta = bpAfter - bpBefore
bpDelta
```

### Calculate the Average Decrease in Blood Pressure

```{r}

meanBpDelta = mean(bpDelta)

meanBpDelta
```

### Identify Which Subjects Had a Decrease in Blood Pressure

```{r}

negBpPatients = names(which(bpDelta < 0))

negBpPatients
```

### Subset Difference Vector to Only Include Negatives

```{r}

bpDeltaNeg = bpDelta[negBpPatients]

bpDeltaNeg

```

### Average Blood Pressure Decrease in Patients Whose Blood Pressure Decreased

```{r}
meanBpDeltaNeg = mean(bpDeltaNeg)

meanBpDeltaNeg
```

## Task 2

### Create Data Frame

```{r}
bpDf = data.frame(patient = names, before = bpBefore, after = bpAfter, delta = bpDelta)

head(bpDf)

```

### Subset Dataframe on Negative

```{r}

bpDf[bpDf$delta < 0, ]

```

### Add Column if Post BP Less than 120:

```{r}

bpDf["bpLow"] = bpDf$after < 120

bpDf
```

### Print Data Frame:

```{r}
knitr::kable(bpDf)
```

## Task 3

### Create Placebo Dataframe

```{r}

bpBeforePlacebo = c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
bpAfterPlacebo = c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)
namesPlacebo = paste("Subject", 1:10, sep = "_")
bpDeltaPlacebo = bpAfterPlacebo - bpBeforePlacebo

bpPlaceboDf = data.frame(patient = namesPlacebo, before = bpBeforePlacebo, after = bpAfterPlacebo, delta = bpDeltaPlacebo)

head(bpPlaceboDf)
```

### Create List with Two Elements (Treatment / Placebo)

```{r}

experiments = list(treatment = bpDf, placebo = bpPlaceboDf)

str(experiments, max.level = 1)
```

### Access First Element Using 3 Different Commands

```{r}
experiments$treatment

```

```{r}
data.frame(experiments[1])
```

```{r}
experiments[[1]]
```

### Access the the "before" column of the "placebo" dataframe:

```{r}

experiments$placebo$before

```

## Task 4

### Add a column in each dataframe called "Status"

```{r}

experiments$treatment$status = character(20)
experiments$placebo$status = character(10)

experiments$treatment$status

```

### Update Status for Treatment

```{r}

for (i in 1:length(experiments$treatment$patient)){
  bpPost = experiments$treatment[i, ]$after
  if (bpPost <= 120){
    experiments$treatment[i, ]$status = "optimal"
  }
  else if (bpPost <= 130){
    experiments$treatment[i, ]$status = "borderline"
  }
  else{
    experiments$treatment[i, ]$status = "high"
  }
  
}

experiments$treatment
```

### Repeat for High

```{r}
for (i in 1:length(experiments$placebo$patient)){
  bpPost = experiments$placebo[i, ]$after
  if (bpPost <= 120){
    experiments$placebo[i, ]$status = "optimal"
  }
  else if (bpPost <= 130){
    experiments$placebo[i, ]$status = "borderline"
  }
  else{
    experiments$placebo[i, ]$status = "high"
  }
  
}

experiments$placebo
```

takes in a list with two data frames in it (a treatment and a placebo data frame) as an argument.
Give no default value.
• takes in an R function (that would find a summary of a numeric column) with the default value
being set to "mean" (notice this is a quoted string).
• Finds the statistic of interest (as defined by the user input) for the pre, post, and diff columns
of both data frames.– Use my_fun \<- get(stat) within the function to get the function from the quoted string.
• These six values should then be returned as a named list with meaningful names- this is a
somewhat challenging part!
4
– I’m going to let you consider what to do but you might create a vector of names that is
created dynamically based on the statistic passed, create a vector with the actual statistic
values, and then assign names() to your vector. Then return that (an atomic vector with
names can be returned instead of a list).
• Finally, apply your function to you list of data frames from previous. Use it without specifying
your statistic, with specifying your statistic as "var", "sd", "min", and "max".

## Task 5

```{r}

get_summary_stat = function(experiment, fun = "mean"){
  
  my_fun = get(fun)
  types = c("treatment", "placebo")
  cols = c("before", "after", "delta")
  statsCollection = c(
    tb = 0,
    ta = 0,
    td = 0,
    pb = 0,
    pa = 0,
    pd = 0
  )
  
  i = 1 ## use this to count up
  for (type in 1:length(types)){
    for (col in 1:length(cols)){
      j = types[type]
      k = cols[col]
      
      statsCollection[i] = my_fun(experiment[[j]][[k]])
      
      text = paste(j , k , fun)
      
      names(statsCollection)[i] = c(text)
      i = i + 1
    }
  }
  
  return (statsCollection)
}

get_summary_stat(experiments)

```

```{r}

print(get_summary_stat(experiments, "sd"))
print(get_summary_stat(experiments, "min"))
print(get_summary_stat(experiments, "max"))
print(get_summary_stat(experiments, "var"))
```
